Collaborative Infrastructure for Test-Driven \\Scientific Model Validation

### 2014-03-15 sciunit-icse14/sciunit-icse14.tex
ICSE14 final

\begin{document}
\conferenceinfo{ICSE}{'14, May 31 - June 7, 2014, Hyderabad, India}
\CopyrightYear{2014}
\crdata{978-1-4503-2768-8/14/05}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Collaborative Infrastructure for Test-Driven\\Scientific Model Validation}

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle
\titleformat{\section}
 {\secfnt}{\thesection}{1em}{\secfnt\MakeUppercase}
 \vspace{10pt}
\section{Introduction}
\begin{figure*}[t]
\vspace{-30px}
\centering
\includegraphics[scale=0.45]{diagram1.pdf}
\vspace{-50px}
\caption{\small Tests are derived from data and models are derived from scientific theories, either directly or through existing data and model repositories. The score table summarizes the performance of a collection of models against a suite of tests. Tests and models are initialized, executed and visualized using \texttt{sciunit} in an IPython notebook stored inside a \emph{suite repository} (e.g. \texttt{saturnsuite}) hosted on a social coding website (e.g. Github). SciDash  discovers and organizes  suite repositories and provides interactive views of the notebooks they contain. Common testing code, capabilities and bridges to external model and data repositories are also collaboratively developed (e.g. \texttt{cosmounit}).}  
\label{fig:sciunit_overview}
\vspace{-10pt}
\end{figure*}

@@

As with many pieces of cyberinfrastructure, early adopters may benefit less than late adopters due to network effects. To incentivize adoption, we look to an increasingly common scientific practice: modeling competitions. Many research communities use these to periodically determine the state of the art, resulting in a form of review paper. Funding sources often include small monetary incentives.
%For example, the quantitative single neuron modeling competition operated over the course of two years  \cite{qsnmc}. 
   SciDash could be used to run these kinds of competitions on a continuous basis atop less \textit{ad hoc} infrastructure (and thus require less effort and central planning from organizers) and we are actively recruiting organizers for such competitions.

@@

\vspace{5px}
\section{Discussion \& Future Directions}\label{discussion}
Instead of attempting to use an existing general-purpose testing infrastructure, we have designed what is essentially a domain-specific language that captures the scientific model validation process in more familiar terms (challenge 6). SciUnit can be used by individual scientists to organize their workflows using IPython notebooks \cite{perez2007ipython} and also within a loosely collaborative workflow, mediated by a social coding tool like Github and a simple portal, SciDash.

@@

While we discuss a simple example based on planetary movement here, we have applied this framework to more realistic problems in neurobiology in collaboration with a large-scale informatics project, OpenSourceBrain, whose developers are enthusiastic about precisely characterizing the scope and validity of the dozens of models they currently maintain (case study in preparation). The tools we describe could also be used for a variety of other scientific disciplines, including within software engineering to validate predictive models of, for example, development time or code reuse.

This paper identifies potential  synergies between software development and model validation practices, and sketches a sociotechnical solution to several challenges to argue that a test-driven development methodology could help improve the scientific process. The scientific model validation process has not yet been extensively studied, so precisely characterizing how scientists attempt to answer the questions in the introduction today, and evaluating how effective tools like the ones we propose are in improving their ability to do so, are likely fruitful avenues for future empirical research.

@@

\section{ACKNOWLEDGEMENTS}
We thank Sharon Crook, Shreejoy Tripathy and Padraig Gleeson for their many helpful discussions, and the anonymous referees and several friends for their useful comments.  
RG was supported in part by NIMH grant R01MH081905. 
The content is solely the responsibility of the authors and does not necessarily represent the official views of the NIH.
\vspace{5px}
% conference papers do not normally have an appendix

>>>>

\begin{document}
\conferenceinfo{ICSE}{'14, May 31 - June 7, 2014, Hyderabad, India}
\CopyrightYear{2014}
\crdata{978-1-4503-2768-8/14/05}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Collaborative Infrastructure for Test-Driven \\Scientific Model Validation}

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
%\IEEEpeerreviewmaketitle
\titleformat{\section}
 {\secfnt}{\thesection}{1em}{\secfnt\MakeUppercase}
 \vspace{10pt}
\section{Introduction}
\begin{figure*}[t]
\vspace{-30px}
\centering
\includegraphics[scale=0.45]{diagram1.pdf}
\vspace{-50px}
\caption{\small Tests are derived from data and models are derived from scientific theories, either directly or through existing data and model repositories. The score table (center) summarizes the performance of a collection of models against a suite of tests. Tests and models are initialized, executed and visualized using \texttt{sciunit} in an IPython notebook stored inside a \emph{suite repository} (e.g. \texttt{saturnsuite}) hosted on a social coding website (e.g. Github). SciDash  discovers and organizes  suite repositories and provides utilities for viewing them online without needing to create a local copy. Logic common across research areas is also collaboratively developed (e.g. \texttt{cosmounit}).}  
\label{fig:sciunit_overview}
\vspace{-10pt}
\end{figure*}

@@

As with many pieces of cyberinfrastructure, early adopters may benefit less than late adopters due to network effects. To incentivize adoption, we look to an increasingly common scientific practice: modeling competitions. Many research communities use these to periodically determine the state of the art, resulting in a form of review paper (e.g. \cite{jolivet_quantitative_2008}). Funding sources often include small monetary incentives.
%For example, the quantitative single neuron modeling competition operated over the course of two years  \cite{qsnmc}. 
   SciDash could be used to run these kinds of competitions on a continuous basis atop less \textit{ad hoc} infrastructure (and thus require less effort and central planning from organizers) and we are actively recruiting organizers for such competitions.

@@

\vspace{5px}
\section{Discussion \& Future Directions}\label{discussion}
Instead of attempting to use an existing general-purpose testing framework (e.g. JUnit or PyUnit), we have designed what is essentially a domain-specific language that captures the scientific model validation process in more familiar terms (in view of challenge 6).% SciUnit can be used by individual scientists to organize their workflows using IPython notebooks \cite{perez2007ipython} and also within a loosely collaborative workflow, mediated by a social coding tool like Github and a simple portal, SciDash.

@@

While we discuss a simple example based on planetary movement here, we have applied this framework to more realistic problems in neurobiology in collaboration with the developers of two large-scale informatics projects, OpenSourceBrain and NeuroElectro, which serve as a source of models and data, respectively. Our tools serve, in essence, as a bridge between these resources. The tools we describe could also be used for a variety of other scientific disciplines, including within software engineering to validate predictive models of, for example, project duration or code reuse.

This work identifies new research directions in software engineering by developing connections between test-driven development \cite{beck2003} and the scientific model validation process, which has not been as extensively studied. Characterizing how scientists today attempt to answer the questions given in the introduction, evaluating how well sociotechnical solutions like the ones we propose here could improve their ability to do so, and studying adoption barriers are likely to be fruitful avenues for future empirical research, and lead to a better understanding of both programmers and scientists.

@@

\section{ACKNOWLEDGEMENTS}
We thank Sharon Crook, Shreejoy Tripathy and Padraig Gleeson for their many helpful discussions, and anonymous referees and several friends for their careful readings.  
RG was supported in part by NIMH grant R01MH081905. 
The content is solely the responsibility of the authors and does not necessarily represent the official views of the NIH.
\vspace{5px}
% conference papers do not normally have an appendix
